Synthesis of Linear Functional Programs

### 2021-07-22 article/article.tex
Refinement types writing

\mypara{Refinement Types} Their syntax is... Their rules are...

\begin{mathpar}
    \infer*[right=(refR)]
    {  }
    {  }
\end{mathpar}

>>>>

\mypara{Refinement Types} Refinement types are defined as a types
with a name and with a predicate (a non-existing predicate is the same as it
being always \emph{true}); dependent types are functions with
refinement types in which the argument type name can be used in the predicates
of the return type (e.g. $x\{\mathsf{Int}\} \lolli y\{\mathsf{Int}\ \vert\ y = x\}$ specifies a function that
takes an Int named $x$ and produces an Int whose value is equal to the input $x$). We extend
the types syntax with our refinement types syntax:
\[
\begin{tabular}{lclc}
    $\tau$ & $\ ::=\ $ & $\dots\ \vert\ \langle\emph{name}\rangle \{\tau\ \vert\ P\}$ \\
    $P$ & $\ ::=\ $ & $P = P\ \vert\ P \neq P\ \vert\ P \vee P\ \vert\ P \wedge
    P\ \vert\ P \Rightarrow P\ \vert\ n = n\ \vert\ n \neq n$ \\
    & & $\vert\ n \leq n\ \vert\ n \geq n\ \vert\ n < n\ \vert\ n > n\ \vert\
    true\ \vert\ false\ \vert\ \langle\emph{name}\rangle$ \\
    $n$ & $\ ::=\ $ & $n * n\ \vert\ n + n\ \vert\ n - n\ \vert\
    \langle\emph{natural}\rangle\ \vert\ \langle\emph{name}\rangle$
    \\
\end{tabular}
\]
The addition of refinement types to the synthesizer doesn't
interfere with the rest of the process. We define the following right and
left rule, to synthesize or consume in synthesis a refinement type, where
$\textsc{getModel}(p)$ is a call to an SMT solver that returns a model of an
uninterpreted function that satisfies
$\forall_{\tilde{1},\tilde{2},\dots,\tilde{n}}\ h_{\tilde{1}} \Rightarrow h_{\tilde{2}} \Rightarrow \dots
\Rightarrow h_{\tilde{n}} \Rightarrow p$, where $\tilde{i}$ is the refinement type
name and $h_{\tilde{i}}$ its predicate for every refinement type in the
propositional contexts; and $\textsc{sat}(p_{\tilde{y}} \Rightarrow p_{\tilde{x}})$ is a call to an SMT solver that
determines universal satisfiability of the implication between predicates (the
left focused proposition subtypes the goal).

\begin{mathpar}
    \infer*[right=(refR)]
    { \textsc{getModel}(p) = M }
    {\Theta/\Theta';\Rho;\Gamma;\Delta \vdash M : \tilde{x}\{A\ \vert\
    p\}\Uparrow }
    \and
    \infer*[right=(refL)]
    { \textsc{sat}(p_{\tilde{y}} \Rightarrow p_{\tilde{x}}) }
    {\Theta/\Theta';\Rho;\Gamma;\Delta; x{:}\tilde{y}\{A\ \vert\ p_{\tilde{y}}\}\Downarrow\ \vdash x : \tilde{x}\{A\ \vert\ p_{\tilde{x}}\} }
\end{mathpar}
